Automatic Identification and Annotation of Tasks in Structured Programs

### 2018-04-09 docs/ACM/paper.tex
filled sections 4.1 4.2 and 4.3

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{images/Absolute_speedups}
\caption{Speedup comparisons -- automatic annotations compare favourably against
manual interventions, and lead to great performance gains.
The Y-axis shows the speedup of either manual intervention, or automatic
annotation (this paper) onto the original programs.}
\label{fig:Absolute_speedups}
\end{center}
\end{figure}

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{images/Opt_Speedups}
\caption{The benefit of our optimizations.
Each optimization (profitability analysis and recursion bounds) lead to
non-trivial speedups.}
\label{fig:Opt_Speedups}
\end{center}
\end{figure}

@@

\begin{table}[h]
\centering
\caption{Different kinds of tasks among the BOTS benchmarks. {\Taskminer}'s versaitily allows the user to define which type of tasks should be mined, annotated or cared for.}
\label{tab:types}
\begin{tabular}{|r|c|c|c|}
\hline
                   & \textbf{Recursive} & \textbf{FunctionCall} & \textbf{Region} \\ \hline
\textbf{fft}       & \checkmark         &                       & \checkmark      \\ \hline
\textbf{fib}       & \checkmark         &                       &                 \\ \hline
\textbf{floorplan} & \checkmark         &                       & \checkmark      \\ \hline
\textbf{sort}      & \checkmark         &                       & \checkmark      \\ \hline
\textbf{sparselu}  &                    & \checkmark            & \checkmark      \\ \hline
\textbf{strassen}  & \checkmark         &                       & \checkmark      \\ \hline
\textbf{uts}       & \checkmark         &                       &                 \\ \hline
\textbf{nqueens}   & \checkmark         &                       & \checkmark      \\ \hline
\textbf{knapsack}  & \checkmark         &                       & \checkmark      \\ \hline
\textbf{health}    & \checkmark         & \checkmark            & \checkmark      \\ \hline
\textbf{jacobi}    &                    &                       &                 \\ \hline
\end{tabular}
\end{table}

%APPLICABILITY
%We'll show TaskMiner application on larger benchmarks, such as those of spec or those of the LLVM's test-suite.
\subsection{Scalability}
\label{sub:scalability}

>>>>

\begin{figure}[h]
\begin{center}
\includegraphics[width=1.1\columnwidth]{images/EVAL1_bots_and_toys.png}
\caption{Speedup comparisons -- automatic annotations compare favourably against
manual interventions, and lead to great performance gains.
The Y-axis shows the speedup of either manual intervention, or automatic
annotation (this paper) onto the original programs.}
\label{fig:Absolute_speedups}
\end{center}
\end{figure}

In order to evaluate {\Taskminer}'s performance, we picked the \textsf{BSC-Bots} along with
five other single source programs\footnote{Toy benchmarks: Boruvka's MST algorithm;
Bellman-Ford's shortest path algorithm; Book filter, an algorithm that performs operations on lines from
a text; Private, the algorithm from Figure \ref{fig:ex_privatize}; and the Depth-first search algorithm.}
and produced four versions for each single source file: the sequential original program; a parallel version
of the program, annotated by hand; a parallel version automatically annotated by {\Taskminer}, without
the task cost model; a parallel version annotated by {\Taskminer} with the cost model enabled. Figure
\ref{fig:Absolute_speedups} show {\Taskminer}'s overall performance compared to the original
sequential and manually annotated programs. For simplification purposes, we address
the difference between the automatic samples in regards to the cost model in section
\ref{sub:optimizations}.

{\Taskminer}'s annotated programs were faster than their sequential counterparts in 13 of the 16
programs. Most of the programs in these benchmarks were classic Divide and Conquer algorithms, such as \textit{Strassen}'s matrix multiplication, \textit{knapsack.c} and \textit{fft.c}.
In 6 of them, the automatically annotated programs were close and sometimes even slightly
above the manually annotated ones. In the programs \textit{fib.c}, \textit{floorplan.c} and
\textit{bellmandford.c}, although the
{\Taskminer}'s version fell behind the sequential ones, it was also faster than the manually annotated
versions, which were clearly slower than the original programs. The manually annotated versions for all
the \textsf{BSC-Bots} benchmarks were obtained from \cite{Duran09}. Regarding the \textit{bellmanford.c}
benchmark, we believe that the slowdown was due to small tasks being dispatched. Even the cost model
wasn't able to pay off for this difference.
In the \textit{fib.c} program, both parallel versions (automatic and manual) were far slower than
the original, specially for big entries. We believe that the compiler we utilized optimized the
classic sequential fibonacci algorithm, while the \textsf{OpenMP} runtime didn't.
Overall, {\Taskminer} has proven to be able to find all the tasks in the \textsf{BSC-Bots} benchmarks and
its annotated versions even surpassed some of the original in performance.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=1\columnwidth]{images/EVAL2_opt.png}
\caption{The benefit of our optimizations.
Each optimization (cost model and recursion bounds) lead to
non-trivial speedups.}
\label{fig:Opt_Speedups}
\end{center}
\end{figure}

In most of the benchmarks, the presence of task cost estimates made little difference in performance of
the automatically annotated programs. In some of them, the cost model was very decisive to 
guarantee a speed up against the sequential versions, though. These can be seen in Figure \ref{fig:Opt_Speedups}.
A coarse gain in speed up was noticed in the \textit{nqueens.c} program, in which the recursion bounding
avoided the creation of smaller and exponentially increasing tasks, which caused a significant speed
up in its execution. 
If on one hand, the cost model is able to avoid the creation and the dispatch of tasks that are too small
and are not going to pay off for the runtime cost, on the other hand computing the cost for tasks can add
some overhead to the programs. This is easily seen in the \textit{bellmanford} program. For larger
matrices, larger is the number of paths to be computed, so in this case, {\Taskminer} will try to
compute the cost for every task, which we believe caused this minor slowdown. However, it's
clear that the cost model greatly improved {\Taskminer}.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=1\columnwidth]{images/EVAL3_test-suite.png}
\caption{{\Taskminer}'s performance in the real world. This chart shows that our algorithm also performs
well on programs that are not deemed inherently parallel, being able to find hidden parallelism
in 20 programs from LLVM's Test Suite.}
\label{fig:testsuite}
\end{center}
\end{figure}

Our tool has also performed well on \textsf{LLVM's Test Suite}, being able to even find hidden parallelism.
From a total of XX single source programs,
{\Taskminer} automatically produced parallel annotated versions of 20 of them, as seen in Figure
\ref{fig:testsuite}. We were able to not only find hidden parallelism in those programs, which are not
considered inherently parallel, but also produce these parallel versions automatically. Amongst
those programs, one caught our attention: \textit{misc/lowercase.c}, which applies a function to every
string in the text. The automatically annotated version of this program was 3.5 times faster than
the sequential one. We have also nice speed up numbers in \textit{misc/ffbench} and \textit{misc/mandel-2}.

\begin{table}[h]
\centering
\caption{Different kinds of tasks among the BOTS benchmarks. {\Taskminer}'s versaitily allows the user to define which type of tasks should be mined, annotated or cared for.}
\label{tab:types}
\begin{tabular}{|r|c|c|c|}
\hline
                   & \textbf{Recursive} & \textbf{FunctionCall} & \textbf{Region} \\ \hline
\textbf{fft}       & \checkmark         &                       & \checkmark      \\ \hline
\textbf{fib}       & \checkmark         &                       &                 \\ \hline
\textbf{floorplan} & \checkmark         &                       & \checkmark      \\ \hline
\textbf{sort}      & \checkmark         &                       & \checkmark      \\ \hline
\textbf{sparselu}  &                    & \checkmark            & \checkmark      \\ \hline
\textbf{strassen}  & \checkmark         &                       & \checkmark      \\ \hline
\textbf{uts}       & \checkmark         &                       &                 \\ \hline
\textbf{nqueens}   & \checkmark         &                       & \checkmark      \\ \hline
\textbf{knapsack}  & \checkmark         &                       & \checkmark      \\ \hline
\textbf{health}    & \checkmark         & \checkmark            & \checkmark      \\ \hline
\textbf{jacobi}    &                    &  \checkmark                     &                 \\ \hline
\end{tabular}
\end{table}

{\Taskminer}'s versatility also arises in regards to the diversity of types of task parallelism that it
can find in a program. Table \ref{tab:types} shows all types of tasks found in our main benchmarks.
There are 3 types of tasks: recursive, which arises from vanes in Divide and Conquer
algorithms; function call, which arises within loops that are not clearly parallel; and region, which 
are similar to function calls, but instead are entire fragments of code within loops that can be executed in
parallel. We provide the user complete decision power to mine for any type of tasks they wish when
running our tool.

%SCALABILITY
\subsection{Scalability}
\label{sub:scalability}
